---
title: "UCINET menu to xucinet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{UCINET menu to xucinet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE)
library(xucinet)
```

Every UCINET menu item the book uses, and the xucinet function that answers it.

The table is generated from the crosswalk in the book repository, not written
out here, so it cannot drift from the list the package and the chapters are both
built against. Whether a routine exists is worked out when this page is built,
so **planned** means planned as of this build and nothing more.

```{r}
csv <- system.file("extdata", "crosswalk-routines.csv", package = "xucinet")
if (!nzchar(csv)) {
  # pkgdown builds articles from the source tree with install = FALSE, where
  # system.file() cannot see inst/. R CMD check builds them from an installed
  # package, where it can. Both have to work.
  csv <- file.path("..", "inst", "extdata", "crosswalk-routines.csv")
}
tbl <- utils::read.csv(csv, stringsAsFactors = FALSE, colClasses = "character")

exists_now <- function(f) {
  f <- sub("[(].*$", "", trimws(f))
  nzchar(f) && grepl("^[a-z_]", f) &&
    exists(f, envir = asNamespace("xucinet"), mode = "function")
}
tbl$status <- ifelse(vapply(tbl$name_2, exists_now, logical(1)), "", "planned")
```

`r sum(tbl$status == "")` of `r nrow(tbl)` routines are written. The rest are
Phase 1 and later; their first-edition names already resolve, and each starts
working the day its routine lands.

```{r, results = "asis"}
for (ch in unique(tbl$chapter)) {
  rows <- tbl[tbl$chapter == ch, ]
  cat("\n## Chapter ", ch, "\n\n", sep = "")
  cat("| 3e | UCINET menu | xucinet | 1e name | |\n")
  cat("|---|---|---|---|---|\n")
  for (i in seq_len(nrow(rows))) {
    r <- rows[i, ]
    cat("| ", r$section,
        " | ", gsub("|", "&#124;", r$menu, fixed = TRUE),
        " | `", r$name_2_raw, "`",
        " | ", if (nzchar(r$name_1e)) paste0("`", r$name_1e, "`") else "",
        " | ", r$status, " |\n", sep = "")
  }
}
```

## Reading this table

The **UCINET menu** column is the path in UCINET 6. A bar inside a path is part
of the path, not an alternation.

The **1e name** column is what *Analyzing Social Networks with R* (1st edition)
called it. Those names still work: they are exported as deprecated wrappers that
name the replacement and forward to it. See `xucinet_1e_names()` for the whole
list, including the several the first edition spelled two ways.

One routine per menu item, and one function per routine. Where UCINET's dialog
has checkboxes, xucinet has arguments: the argument vocabulary is fixed across
the package, so learning three functions teaches you the rest.
