---
title: "Getting started, for UCINET users"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started, for UCINET users}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(xucinet)
# Loaded explicitly: lazy data is attached when the package is installed, but
# not when pkgdown builds this from the source tree.
data(campnet, package = "xucinet")
```

This is a stub. It covers what works today — Phase 0, which is file handling,
the network object and one pilot routine — and grows a section per phase as the
routines land.

## Nothing to learn but R punctuation

Everything you type is lowercase: function names, argument names, argument
values, dataset names. There is no shift key anywhere in the interface, which
kills the commonest beginner error. Type `x` and press Tab and the package
unfolds.

```{r}
xdensity(campnet)
```

That is the whole idea: the report is UCINET's, laid out the way UCINET lays it
out, and the object behind it is an ordinary R structure you can compute on.

```{r}
res <- xdensity(campnet)
res$summary$Density
```

## Your data, whatever shape it is in

Every routine takes a network as its first argument, and a network can be a
matrix, a data frame, an igraph object, a statnet `network`, or a file name.

```{r}
m <- matrix(c(0,1,1, 1,0,0, 1,0,0), 3, 3)
xdensity(m)
```

A dataset name works too, as it does on UCINET's command line — no `data()`
first:

```{r}
xdensity("campnet")
```

## Reading and writing UCINET's own files

`xread()` detects the container from the file and the layout from the shape of
what is inside it, so most of the time there is nothing to say beyond the name.

| format | read | write |
|---|---|---|
| `##h`/`##d`, all six header versions | `xreaducinet()` | `xsaveucinet()` |
| `.uci`, the single-file JSON | `xreaduci()` | `xsaveuci()` |
| DL | `xreaddl()` | `xsavedl()` |
| VNA | `xreadvna()` | `xsavevna()` |
| csv, Excel | `xread()` | `xsave()` |

```{r}
f <- system.file("goldens", "ucinet", "davis.##h", package = "xucinet")
davis_from_file <- xreaducinet(f)
dim(davis_from_file)
```

## Finding the routine you want

`xhelp()` searches on the UCINET menu path, the measure name, plain English, or
the name the first edition used.

```{r}
xhelp("betweenness")
```

Routines still to be written are listed and marked `planned`, so you can see
what is coming and what it will be called.

## Where things are

- Every dataset the book uses ships with the package: `data(package = "xucinet")`.
- The full menu-to-function map is in `vignette("crosswalk")`.
- Every deliberate difference from UCINET is in `vignette("differences")`.
- First-edition names still work, with a message pointing at the new one.
